\section{Overview}
\label{sec:overview}

When an AI agent needs to do a task, it must know what tools or capabilities it has available. Right now, the most common way to do this is to write a list of all available tools and paste the whole list into the AI's instructions at the start of every task. This is called \textit{tool use}. It is simple to set up, and it works when the number of tools is small.

The problem appears as soon as the list gets longer. Research has shown that when an AI model is given a long list of tools to read at the beginning, it gets confused \cite{patil2023gorilla}. It starts making mistakes, sometimes trying to use tools that do not even exist, or ignoring the correct tool completely. This is because the long list takes up space in the AI's memory that should be used for thinking about the actual task \cite{shi2024context}. The more tools in the list, the worse the problem gets.

A newer approach is now being used in autonomous agent frameworks like OpenClaw \cite{openclaw_2025}. Instead of giving the AI the full list of tools all at once, it only shows the AI the names and short descriptions first. When the AI decides it needs a specific tool, only then does it load the full instructions for that tool. Once the tool is finished being used, those instructions are removed from memory, and the AI goes back to working with just the short list. This step-by-step loading pattern is called \textit{Progressive Disclosure}.

The idea behind progressive disclosure is simple: show the AI only what it needs, when it needs it. This should help the AI stay focused, make fewer mistakes, and use fewer tokens per task --- which also reduces cost. OpenClaw \cite{openclaw_2025} already implements this approach through its \texttt{SKILL.md} file system, where each capability is stored as a separate structured file that is loaded on demand.

However, nobody has yet done a proper test to find out whether progressive disclosure actually performs better than the traditional flat tool list. The claims made by framework developers have not been independently measured or tested side by side. This proposal describes a study that will carry out that test in a clear and reproducible way, using OpenClaw as the agent framework and both local AI models run through Ollama \cite{ollama_2024} and cloud-based models --- Gemini 3.1 Pro by Google \cite{google_gemini_2025} as the primary cloud model and GPT-4o by OpenAI \cite{openai_gpt4o_2024} as a second cloud model. The goal is to give researchers and developers honest numbers so they can make better decisions about how to build AI agents.

\newpage
