\section{Background and General Goals}
\label{sec:background}

\subsection{Background}

An AI agent is a program that uses a language model to think through a task and take actions to complete it. For example, an agent might be told to read a PDF, extract specific information from it, search the web to verify a detail, and then write the result into a file. Unlike a regular chatbot, an agent does not just reply with text. It actually carries out steps, uses tools, and adapts when something goes wrong \cite{yao2022react}.

To do any of this, the agent needs to know what tools are available to it. The standard way to handle this is called \textit{tool use}. The developer writes out all the available tools --- their names, what they do, and what inputs they need --- and pastes this entire list into the agent's instructions at the very beginning of every task. The agent reads through the list and picks the tool it needs \cite{schick2023toolformer}.

This works well when there are only a few tools. The problem comes when the number of tools grows. AI models have a limited amount of working memory, called a context window. When a long list of tool descriptions takes up a large portion of that memory, there is less space left for actually thinking about the task \cite{shi2024context}. Research has confirmed that as more tools are added to the list, the AI makes more mistakes --- including trying to call tools that do not exist \cite{patil2023gorilla}. This issue is even worse for smaller, locally-run models that have shorter context windows to begin with.

A different approach is now being used in the field. Instead of loading all tool descriptions at once, some frameworks store each capability as a separate file. The agent only sees a short list of tool names and one-line descriptions to start with. When it decides it needs a specific tool, it loads the full instructions for that tool at that moment. Once the tool is done, those instructions are unloaded. This step-by-step approach is called \textit{Progressive Disclosure} and it is built into the OpenClaw autonomous agent framework \cite{openclaw_2025} through its \texttt{SKILL.md} file system.

The problem is that nobody has done a proper, honest test to find out whether progressive disclosure actually works better. The developers of OpenClaw claim it is more efficient, but that claim has never been independently tested. This thesis will carry out that test.

\subsection{Statement of Objectives}

This thesis will do three things:

\begin{enumerate}
    \item Test how accurately an AI agent finishes tasks when all tools are loaded at once (the standard way) versus when tools are loaded one at a time as needed (progressive disclosure via OpenClaw \cite{openclaw_2025}). The test will be run using two local models --- \texttt{gemma4:4b} by Google DeepMind \cite{gemma4_2026} and \texttt{qwen3.5:4b} by Alibaba \cite{qwen3_2025}, both run through Ollama \cite{ollama_2024} --- and two cloud models --- Gemini 3.1 Pro by Google \cite{google_gemini_2025} as the primary and GPT-4o by OpenAI \cite{openai_gpt4o_2024} as a second cloud reference.

    \item Measure how many tokens are used per task under each approach, and how that changes as the number of available tools grows from 10 to 20. Token count is important because it directly determines cost for cloud models and speed for local models.

    \item Count how many times the agent tries to call a tool that does not exist (a \textit{hallucinated tool call}) under each approach, and check whether the agent follows the step-by-step instructions inside a skill file correctly (called the \textit{trajectory compliance score}).
\end{enumerate}

\subsection{Significance of the Project}

Developers building AI agents today face a real choice: load all tools at once and risk confusion, or use a structured skill system and risk complexity. Right now, there is no independent evidence showing which approach actually leads to better results. This thesis will fill that gap by running a controlled, repeatable experiment with real tools, real tasks, and real measurements. The results will help developers and researchers make better decisions based on actual data rather than assumptions or marketing claims.
