\section{Overview}
\label{sec:overview}

Every day, many organizations deal with large numbers of documents like invoices, registration forms, and official records. Someone needs to read each one and pull out specific pieces of information, such as a total amount, a name, or a date. Doing this by hand is very slow and expensive. So many organizations now use AI to do it automatically. This is called \textit{data extraction}.

Right now, most organizations do this by sending their documents to an online AI service. The service reads the documents and sends back the information, but it charges a small fee every time. When you have thousands of documents every day, those fees add up very fast. There is also a second problem: private documents have to leave the organization and go to someone else's server, which raises serious privacy and security concerns.

A newer option is now available. Small AI models can now run entirely on a laptop with no internet connection, managed through a free tool called Ollama \cite{ollama_2024} which handles downloading and running them automatically. Three such models will be tested in this study: \texttt{deepseek-ocr} (DeepSeek AI) \cite{deepseek_ocr_2024}, \texttt{gemma4:e4b} (Google DeepMind) \cite{gemma4_2026}, and \texttt{qwen3.5:4b} (Alibaba) \cite{qwen3_2025}. A fourth model, \texttt{Phi-4-multimodal-instruct} (Microsoft) \cite{phi4mini_2025}, will also be tested if hardware resources allow. These models are free to use, keep all data private, and can read and understand document images without sending anything to the internet. However, organizations do not yet have clear, honest numbers that show whether these local models are accurate enough, fast enough, or cheap enough compared to paid online options. There is also an open question about whether the results change depending on how the instruction given to the AI is written.

This proposal describes a study that will test and compare these two approaches side by side. On one side are three local AI models run through Ollama; on the other are two paid online AI services (GPT-4o by OpenAI \cite{openai_gpt4o_2024} and Gemini 3.1 Pro by Google \cite{google_gemini_2025}). The comparison covers three things: how accurate each one is (tested using different ways of writing the instruction), how fast each one is, and how much each one costs. The goal is to give organizations clear numbers so they can decide which option is right for them.
